Test if analytic derivation of geodesic distance for curves from sim_functional_data.R

### 2019-06-20 index.Rmd
scms now uses silverman's rule but keeps isotropic kernel bandwidth matrix using sigma_min of cosmic reconstruction paper.
added scenario 4

@@

## Comparison to other methods

@@

We compare our method to the following ones:

* **RD** Apply IsoGeo on the raw data $Y_1,\ldots,Y_n$ to obtain an estimator $\hat G_{\textrm{RD}}$. Note that this procedure can only be used if the grid $T_i$ is the same for each $i$.
* **SS** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing. Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.
* **pI** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing. Calculate the weighted graph of step I of Isomap with a penalty (describe the penalty of Chen and Muller). Apply step II of Isomap to obtain $\hat G_{\textrm{pI}}$.
* **RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.

# Assessment of geodesic distance estimates
We introduce a new tool to assess pairwise geodesic distance estimation. This is the ROC curve with $\epsilon$ on the $x$-axis and degree to which near-$\epsilon$ isometry holds, i.e.\ the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.

>>>>

## Abstract

Title: Pairwise geodesic distance estimation for functional data residing near a nonlinear manifold.

Abstract: Certain functional data of interest reside in non-linear spaces; the class of probability density functions provides one such example. Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration and treatment. In this work we set ourselves the task of calculating the geodesic distance between two functions living near a non-linear manifold due to noise. This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold. Good estimation of the geodesic distance has implications for many downstream tasks in functional data analysis, including functional clustering, and nonparametric functional regression.

@@

## Simulaton Study

@@

# Comparison to other methods

* **(Raw Data) RD** Apply IsoGeo on the raw data $Y_1,\ldots,Y_n$ to obtain an estimator $\hat G_{\textrm{RD}}$. Note that this procedure can only be used if the grid $T_i$ is the same for each $i$.
* **(Spline Smoothing) SS** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing. Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.
* **(Penalized-Isomap) pI** Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing. Calculate the weighted graph of step I of Isomap with a penalty (describe the penalty of Chen and Muller). Apply step II of Isomap to obtain $\hat G_{\textrm{pI}}$.
* **(Random Projection) RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.

# Assessment of geodesic distance estimates
We introduce a new metric for assessing the quality of a pairwise geodesic distance estimator. This is the ROC curve with $\epsilon$ on the $x$-axis and degree to which near-$\epsilon$ isometry holds, i.e.\ the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.

### 2019-06-20 test_analytic_geodesic_distance.Rmd
scms now uses silverman's rule but keeps isotropic kernel bandwidth matrix using sigma_min of cosmic reconstruction paper.
added scenario 4

```{r}
library(reticulate)
# sometimes reticulate requires specific instruction on where to find your python
# first call Sys.which("python") and then copy that path into

use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)
#use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
```

@@

The function sim_functional_data.R allows specification of several simulatin scenarios. The samplesize is purposefully set very high for reasons explained below.

```{r}
# Generate data
sim <- sim_functional_data(sce=3,samplesize=100)
```

@@

If the analytic derviation is correct, then estim_geo_noiseless_data should be extremely similar to analytic_geo. We can assess their closeness using assess_goodness_estimation.R. Running this notebook for sce=1 gives rmse on the order of e-16.
```{r}
library(MESS)
source('assess_goodness_estimation.R')
assess_goodness_estimation(estim_geo_noiseless_data,sim$analytic_geo)
```

>>>>

```{r}
library(reticulate)

use_python('/usr/local/bin/python3.7',required=TRUE)
# use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
# use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)

```

Check with py_config()

```{r}
py_config()
```

@@

The function sim_functional_data.R allows specification of several simulatin scenarios. The samplesize should be set very high so that Floyd's algorithm works well on the noiseless data. The number of time grid points K also should be high so that the vector data really represents the functional curve.

```{r}
# Generate data
sim <- sim_functional_data(sce=4,samplesize=1500,K=150,plot_true=1)
```

@@

If the analytic derviation is correct, then estim_geo_noiseless_data should be extremely similar to analytic_geo. We can assess their closeness using assess_goodness_estimation.R. Three assessment metrics are returned. If the calculation is correct, rmse should be close to zero and epsilon_isometry_auc and pearson_corr should both be close to 1. Note that running this notebook for sce=1 gives rmse on the order of e-16.
```{r}
library(MESS)
source('assess_goodness_estimation.R')
assess_goodness_estimation(estim_geo_noiseless_data,sim$analytic_geo)
```
